# Analysis of longitudinal data

```{r}
#libraries needed
library(ggplot2)
library(dplyr)
library(tidyr)
library(lme4)

```

I read the BPRS and RATS (wide form) and BPRSL and RATSL (long form) and check if they are ok.

```{r}
BPRS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/BPRS.txt", sep = " ", header = T)
RATS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt", sep = "\t", header = TRUE)

BPRSL <- read.csv("BPRSL.csv")
RATSL <- read.csv("RATSL.csv")
glimpse(BPRSL)
glimpse(RATSL)

BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)
BPRSL$treatment <- factor(BPRSL$treatment)
BPRSL$subject <- factor(BPRSL$subject)
glimpse(BPRSL)
glimpse(BPRS)

RATS$ID <- factor(RATS$ID)
RATS$Group <- factor(RATS$Group)
RATSL$ID <- factor(RATSL$ID)
RATSL$Group <- factor(RATSL$Group)
glimpse(RATSL)
glimpse(BPRS)
```

**1** Analysis of the RATS data.

Below you can see individual weight gaining profiles by diet group for the RATS data.
 
```{r}
ggplot(RATSL, aes(x = Time, y = Weight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(RATSL$Weight), max(RATSL$Weight)))
```

Almost all the rats are gaining weight during the study. The rats in group one are *very* much lighter throughout the study, also in the beginnig. In every group, there is one rat, which differs from the other rats (an outlier).

Next, I will standardize the variable Weight and plot again after standardization.

```{r}
RATSL <- RATSL %>%
  group_by(Time) %>%
  mutate(stdWeight = (Weight - mean(Weight))/sd(Weight) ) %>%
  ungroup()

ggplot(RATSL, aes(x = Time, y = stdWeight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(name = "Standardized Weight")

```

The next step is to do the summary measure analysis of RATSL data

```{r}
#number of individuals in group 1 is **8** and in group 2 and 3 **4**
#I don't know how to deal with that.

n <- 4

RATSS <- RATSL %>%
  group_by(Group, Time) %>%
  summarise( mean = mean(Weight), se = sd(Weight)/sqrt(n) ) %>%
  ungroup()

ggplot(RATSS, aes(x = Time, y = mean, linetype = Group, shape = Group)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2,3)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.8,0.8,0.8)) +
  scale_y_continuous(name = "mean(Weight) +/- se(Weight)")

```

Above you can see the mean response profiles for the three diet groups in the RATS data. Unfortunately, there is something wrong with this figure.

Let's try something else.

```{r}
# I create a summary data by diet group and subject(ID) with mean as the summary variable

RATSL10S <- RATSL %>%
  filter(Time > 1) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()

glimpse(RATSL10S)

#...and draw a boxplot of the mean versus diet group

ggplot(RATSL10S, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(Weight), days 1-64")
```

Above you can see boxplots of mean summary measures for the three diet groups. There is an outlier in every group, as earlier seen.

**Summary:** The rats (n=8) in the group 1 are much smaller than the rats in the group 2 (n=4) and 3 (n=4). Almost all the rats gained weight during the study. There's an outlier in each group.


**2** Analysis of the BPRS data.

```{r}
ggplot(BPRSL, aes(x = week, y = bprs, group = subject)) + 
  geom_line() +
  scale_x_continuous(name = "Time (weeks)", breaks = seq(0, 8, 2))+
    
  scale_y_continuous(name = "BPRS (points)") +
  theme(legend.position = "top")

ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
  
```

Below are the results from fitting *a linear regression model* to BPRS data with BPRS-points as response variable and treatment and week as explanatory variables. 

```{r}
BPRS_reg <- lm(bprs ~ week + treatment, data = BPRSL)
summary(BPRS_reg)
```

And the next step, below the results from fitting *a random intercept model* to BPRS data with week and treatment as explanatory variables.

```{r}
BPRS_ref <- lmer(bprs ~ week + treatment + (1 | subject), data = BPRSL, REML = FALSE)
summary(BPRS_ref)
```

Below *a random intercept and random slope model* and *an ANOVA test* on the two models.

```{r}
BPRS_ref1 <- lmer(bprs ~ week + treatment + (week | subject), data = BPRSL, REML = FALSE)
summary(BPRS_ref1)
anova(BPRS_ref1, BPRS_ref)
```

p-value of the likelihood ratio test is 0.02636. The value is low meaning a good fit against the comparison model.


Below *a random intercept and random slope model with the interaction* and *an ANOVA test* on the two latest models.

```{r}
BPRS_ref2 <- lmer(bprs ~ week + treatment + (week | subject) + week * treatment, data = BPRSL, REML = FALSE)
summary(BPRS_ref2)
anova(BPRS_ref2, BPRS_ref1)
```

The next step is to create *fitted BPRS profiles from the interaction model*.

```{r}
Fitted <- fitted(BPRS_ref2)
BPRSL <- BPRSL %>%
  mutate(Fitted)

ggplot(BPRSL, aes(x = week, y = Fitted, group = subject)) +
  geom_line() +
  scale_x_continuous(name = "Time (weeks)", breaks = seq(0, 8, 2)) +
  scale_y_continuous(name = "Fitted BRPS (points)") +
  theme(legend.position = "top")

ggplot(BPRSL, aes(x = week, y = Fitted, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
```

**Summary:** During the study BPRSL points decreased in both groups. There were no significant difference between treatments.