# Regression and model validation


**1.** Data of this study includes basic information about the students of a statistic course, their attitude, different kind of learning types and academic success. Data are available from 166 students. Questions considering different learning types have been combined into combination variables called deep, surface and strategic learning. Other variables are gender, age, attitude and exam points.


```{r}
lrn14 <- read.csv("learning2014.csv")
library(dplyr)
keep_columns <- c("gender","Age","Attitude", "deep", "stra", "surf", "Points")
learning2014 <- select(lrn14, one_of(keep_columns))
str(learning2014)
dim(learning2014)
```


**2.** Below you can see the summaries of the variables and a graphical overview of the data. There are more female participants in this data. Age or gender didn't affect the points. There's a correlation between points and attitude.


```{r}
summary(learning2014)

library(GGally)
p <- ggpairs(learning2014, mapping = aes(col = gender), lower = list(combo = wrap("facethist", bins = 20)))
p
```


**3. and 4.** I did choose three explanatory variables (attitude, strategic learning and surface learning) and fit the in a regression model where exam points are the target variable. Summary shown below.


```{r}
my_model2 <- lm(Points ~ Attitude + stra + surf, data = learning2014)
summary(my_model2)
```


While strategic and surface learning didn't have a statistically significant relationship with points, I also tested deep learning and age as explanatory variables. Summary shown below.


```{r}
my_model2 <- lm(Points ~ Attitude + deep + Age, data = learning2014)
summary(my_model2)
```


In a model in which I used attitude and strategic learning as explanatory variables (summary below), attitude has a statistical significance p <0.001 and strategic learning p < 0.1. Multiple R-squared is 
0.2048 meaning that about 20 % of the variation in exam points is explained by the variation in attitude and strategic learning.


```{r}
my_model2 <- lm(Points ~ Attitude + stra, data = learning2014)
summary(my_model2)
```


I also tested simple regression model (attitude vs points), where p < 0.001. Multiple R-squared is 
0.1906 meaning that a bit less that 20 % of the variation in exam points is explained by the variation in attitude only.


```{r}
qplot(Attitude, Points, data = learning2014) + geom_smooth(method = "lm")
my_model <- lm(Points ~ Attitude, data = learning2014)
summary(my_model)
```


**5.** I produced diagnostic plot for simple regression model (attitude vs points). 

QQ-plot: both ends of the line curve, errors of the model are not normally distributed

Residuals vs Fitted values: no pattern, errors don't depend on the explanatory variable

Residuals vs Leverage: no observation has unusually high impact 
 
 
```{r}
plot(my_model, which = c(1,2,5))
```

Here we go again...
