# Dimensionality reduction techniques

The data "human" are described in https://github.com/EmmiD/IODS-project in the file create_humanweek5.R.

```{r}
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", sep= ",", row.names = 1)

#libraries needed
library(GGally)
library(corrplot)
library(ggplot2)
library(dplyr)
library(tidyr)
library(FactoMineR)

```

**1.** A graphical overview of the data and summaries of the variables are shown below.

```{r}
summary(human)
ggpairs(human)

cor_matrix<-cor(human) %>% round(digits = 2)
corrplot(cor_matrix, method="circle", type="upper", cl.pos="b", tl.pos="d", tl.cex = 0.6)
```

Mean female-male rate of reaching secondary schooling is 0.85 and mean female-male rate at labour force is 0.70 pointing out the differences between the social status between female and male worldwide. In many variables it's possible to see huge differences between min and max values, for example maternal mortality ratio aand expected years of education.

We can see a strong positive correlation specially between *life.exp* and *edu.exp*, and also between *life.exp* and *GNI*, *life.exp.* and *edu2.fm*, *GNI* and *edu.exp*, *GNI* and *edu2.FM*, *mat.mor* and *ado.birth*, and *edu2.FM* and *edu.exp.*

We can see a strong negative correlation specially between *mat.mor* and *life.exp*, but also between *mat.mor* and *GNI*, *mat.mor* and *edu.exp*, *mat.mor* and *edu2.FM*, *ado.birth* and *GNI*, *ado.birth* and *life.exp*, *ado.birth*, and *edu.exp* and *ado.birth* and *edu2.FM*.

None of these correlations is surprising.

**2** PCA on the not standardized data is shown below. The first principal component captures all the variability.

```{r}
pca_human <- prcomp(human)
biplot(pca_human, choices = 1:2, cex = c(0.5, 1))
s <- summary(pca_human)
pca_pr <- round(100*s$importance[2,], digits = 1)
pca_pr
```

**3** PCA on the standardized data shown below. The first principal component captures 53,6 % of the variability, the second 16,2 %.

```{r}
human_std <- scale(human)
pca_human <- prcomp(human_std)
biplot(pca_human, choices = 1:2, cex = c(0.4, 1))
s <- summary(pca_human)
pca_pr <- round(100*s$importance[2,], digits = 1)
pca_pr
```

Result differ a lot. The problem with unstandardized variables is that the variances are very different. That's why the variables with the largest variances overrun the PCA results.

**4** We can see, that female labor force participation and female in the parliament have surprisingly low correlation with other variables. These two variables are mainly contributing to the second principal component.

Adolescent births and maternal mortality are highly positively correlated, as expected. Also schooling, gross national income and life expectancy have a strong positive correlation. These also have a negative correlation with adolescent births and maternal mortality. These 6 variables are mainly contributing to the first principal component.

**5** Exploring the tea data:

```{r}
data(tea)
str(tea)
dim(tea)
summary(tea)
gather(tea) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()

```

Multiple correspondence analysis of the tea data
```{r}
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")
tea_time <- select(tea, one_of(keep_columns))

mca <- MCA(tea_time, graph = FALSE)
summary(mca)
plot(mca, invisible=c("ind"), habillage = "quali", graph.type = "classic")
```

The first dimension captures 15 % of variance and the second 14 %. There is a strong link between the first dimension and variables *how* and *where*.

*Unpackaged* and *tea shop* differ most from all the other variables.