# Logistic regression


```{r}
#libraries needed
library(dplyr)
library(ggplot2)
library(tidyr)
library(gmodels)
```

Data of this study are combined from two datasets: one regarding the performance in mathematics and other in Portuguese language. Data is collected from two Portuguese secondary schools. It includes student grades, social and school related features together with alcohol consumption.

```{r}
pormath <- read.table("https://github.com/rsund/IODS-project/raw/master/data/alc.csv", sep= ",", header=TRUE)

colnames(pormath)
```
My personal hypothesis is that male and students who go out with their friends more often use more alcohol than female and those student who stay at home (sex, goout). I also think that student who want to educate higher use less alcohol (higher) and students who has more absences use more alcohol (absences).

```{r}

gather(pormath) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()


#boxplot high use and goout
g2 <- ggplot(pormath, aes(x = high_use, y = goout, col = sex))
g2 + geom_boxplot()

#boxplot high use and absences
g4 <- ggplot(pormath, aes(x = high_use, y = absences, col = sex))
g4 + geom_boxplot()

#crosstable high use and sex
CrossTable(pormath$high_use, pormath$sex)

#crosstable high use and "higher"
CrossTable(pormath$high_use, pormath$higher)

```

From the boxplots I can see that students who go often out, are significantly more often high users of alcohol, specially men. Same trend can be seen in absences. In crosstabs high_use and sex, we can see that OR is 2,5 in male compared to female.

KESKEN, NOT FINISHED